Our Title

### 2015-05-05 draft.tex
Merged in comments for the Domestic Portal and Border Monitoring section.

As with any foreign terrorist threat, it is imperative to assess the United States' domestic capabilities for the detection and interception of weapons and materials that pose a significant risk to American security. In the context of this section, we specifically consider nuclear weapons and materials that have the potential to be detonated domestically after being smuggled past American borders. We examine current and past efforts and capabilities to prevent such weapons from entering the United States undetected, as well as the possibilities for the development and implementation of better detection technologies.

Two of the relevant governmental organizations regulating domestic borders and security are the Federal Bureau of Investigation (FBI) and the Department of Homeland Security (DHS). Together, these organizations are responsible for the security of the United States against both domestic and foreign threats. They coordinate intelligence gathering as well as threat-response and eventually coordinate with the justice system. With these goals in mind, a significant portion of their work deals with the monitoring of borders. The Department of Homeland Security alone commands more than 180,000 personnel, operating such sub-groups as the U.S. Coast Guard, U.S. Customs and Border Protection (CBP), and the Immigration and Customs Enforcement (ICE) \cite{Robb2005}. One of the most relevant DHS sub-groups is the Domestic Nuclear Detection Office (DNDO). The DNDO is responsible for the development and deployment of radiation and detection equipment to assist any other federal agency, including the other DHS sub-organizations and the FBI \cite{UnitedStatesGovernmentAccountabilityOffice2013}.

The aforementioned government agencies have correctly addressed border and seaport monitoring as a location of high potential for the interception and detention of nuclear weapons and materials. In 2005, the CBP in collaboration with the American Association of Port Authorities (AAPA) had already begun deployment of a large network of Radiation Portal Monitors (RPMs) with a comprehensive plan for 100\% screening of the locations seen in \autoref{fig:RPM_map} \cite{Simmons2005}.

@@

The CPB/AAPA RPM plan outlines a phased implementation targeting locations on a priority basis around the U.S. The plan additionally details a technology collaboration with the Pacific Northwest National Lab (PNNL), who was responsible for the development of cutting-edge, harmless and passive detection systems (illustrated in \autoref{fig:RPM_designs}). In 2005, RPM systems had already been deployed at 14 terminals, with another 18 in construction and 27 in final design \cite{Simmons2005}.

@@

Eight years after the initial development of the RPM plan, the Office of the Inspector General published a report documenting the progress made in the RPM development specifically at seaports \cite{DepartmentofHomelandSecurityDHS2013}. The document reports 444 active RPMs screening 99\% of inbound containerized cargo at large seaports, with the other 1\% of cargo entering low-volume seaports (this 1\% presumably not monitored).

@@

In order for the U.S. to be effectively guarded against potential nuclear terrorist threats from cunning organizations such as ISIL, a streamlined, more efficient approach must be taken to portal and border monitoring. True, the initial development and deployment of RPM technology seems to have been fairly effective based on publicly - available reports \cite{Simmons2005,DepartmentofHomelandSecurityDHS2013}. The caveat to these programs is their reported deterioration over a decade of dormancy.

@@

This document suggests the aggressive continued advancement of such programs. The continued development of new technologies, as outlined in \autoref{}, must be exploited for every feasible advancement leading to total success in national security. This document suggests a priority-phased plan outlining a strong technological collaboration with government labs, similar to the plan put forth in 2005, in order to develop and deploy the next generation of RPMs \cite{Simmons2005}. Such a plan would be managed by the DNDO and executed by the CPB and AAPA. This plan, however, can only be effective given the successfully executed deployment of any resulting technology. Technological failures such as the ASP development program are acceptable so long as a workable solution is subsequently achieved.

>>>>

As with any terrorist threat, it is imperative to assess the United States' capacity for the detection and interception of weapons and materials that pose a significant risk to American security. In the context of this section, we specifically consider nuclear weapons and materials that have the potential to be detonated domestically after being smuggled into America. We examine current and past efforts and capabilities to prevent such weapons from entering the United States undetected, as well as the possibilities for the development and implementation of better detection technologies.

Two of the relevant governmental organizations regulating domestic borders and security are the Federal Bureau of Investigation (FBI) and the Department of Homeland Security (DHS). Part of the responsibility of these organizations is ensuring the security of the United States from both domestic and foreign threats. They coordinate intelligence gathering, threat-response, and, eventually, coordinate with the justice system. With these goals in mind, a significant portion of their work deals with the monitoring of borders. The Department of Homeland Security alone commands more than 180,000 personnel, operating such sub-groups as the U.S. Coast Guard, U.S. Customs and Border Protection (CBP), and the Immigration and Customs Enforcement (ICE) \cite{Robb2005}. One of the most relevant DHS sub-groups is the Domestic Nuclear Detection Office (DNDO). The DNDO is responsible for the development and deployment of radiation and detection equipment to assist any other federal agency, including the other DHS sub-organizations and the FBI \cite{UnitedStatesGovernmentAccountabilityOffice2013}.

The aforementioned government agencies have correctly assessed border and seaport points of entry as a high risk location for the interception and detention of nuclear weapons and materials. In 2005, the CBP in collaboration with the American Association of Port Authorities (AAPA) had already begun deployment of a large network of Radiation Portal Monitors (RPMs) with a comprehensive plan for 100\% screening of the locations seen in \autoref{fig:RPM_map} \cite{Simmons2005}.

@@

The CPB/AAPA RPM plan outlines a phased implementation approach targeting locations around the U.S. based on their assessed risk. The plan additionally details a technology collaboration with the Pacific Northwest National Lab (PNNL), which is responsible for the development of cutting-edge, harmless and passive detection systems (illustrated in \autoref{fig:RPM_designs}). In 2005, RPM systems had already been deployed at 14 terminals, with another 18 in construction and 27 in final design \cite{Simmons2005}.

@@

Eight years after the initial development of the RPM plan, the Office of the Inspector General published a report documenting the progress made in the RPM development specifically at seaports \cite{DepartmentofHomelandSecurityDHS2013}. The document reports 444 active RPMs screening 99\% of inbound containerized cargo at large seaports, with the other 1\% of cargo entering low-volume seaports (this 1\% is presumably not monitored).

@@

In order for the U.S. to be effectively guarded against potential nuclear threats from non-state actors such as ISIL, a streamlined approach must be taken to portal and border monitoring. While the initial development and deployment of RPM technology seems to have been fairly effective based on unclassified reports, the caveat to these programs is their reported deterioration over a decade of dormancy \cite{Simmons2005,DepartmentofHomelandSecurityDHS2013}.

@@

This document suggests the aggressive continued advancement of such programs. The continued development of new technologies, as outlined in \autoref{}, must be exploited for every feasible advancement leading to total success in national security. This document suggests a priority-phased plan outlining a strong technological collaboration with government labs, similar to the plan put forth in 2005, in order to develop and deploy the next generation of RPMs \cite{Simmons2005}. Such a plan would be managed by the DNDO and executed by the CPB and AAPA. This plan, however, can only be effective given the successfully executed deployment of any developing technology. Technological failures such as the ASP development program are acceptable so long as workable alternatives exist.
